---
title: "data_check_update"
author: JoJo Pikowski
date: 2025/02/02
format: pdf
---

```{r}
#libraries
library(dplyr)

```

```{r}
head(events_full)
```

```{r}
summary(events_full)

```

## Overall counts -- no ER

```{r}
#unique PID values
unique_pid_count <- events_full %>%
  summarise(unique_pids = n_distinct(pid))

unique_pid_count

```


```{r}
#num of events seperated by type
event_type_count <- events_full %>%
  group_by(event_type) %>%
  summarise(count = n())

event_type_count

```

```{r}
#median of coverage days
coverage_duration_stats <- events_full %>%
  summarise(mean_coverage_duration = mean(coverage_duration_days, na.rm = TRUE),
            median_coverage_duration = median(coverage_duration_days, na.rm = TRUE))

coverage_duration_stats

```

```{r}
#num of occurrences of conditions
condition_counts <- events_full %>%
  select(alzheimers:non_ibis_observation_admission_days) %>%
  summarise_all(~ sum(!is.na(.)))

condition_counts
```

# UniCare Study Participants (Pre-Treatment) 
- purple is post ibis pretreatment is preibis

```{r}


```


## UniCare Study Participants (In Treatment)
```{r}


```


## All Ibis Member Average (Pre-Treatment)
```{r}


```


## All Ibis Member Average (In Treatment)
- between ibis_cov_start and ibis_cov_end
```{r}


```

## Look at admissions post-ibis
```{r}

```


# How many ibis_cov_start date = 0 and then change to the start date or a high number --- all in the pre period

- count total admissions pre and post for both the unicare-study and the whole group

- compare days: number of days total of coverage for unicare-study vs everybody pre and post treatment

when you add his column of preibis in patient it should what we get from the table 

- add all the coverage days and also do counts by patient 
ultimately find the inpatient events for each patient
