---
title: "Unicare Pre vs Post Analysis"
format: html
editor_options:
  chunk_output_type: console
---

```{r knitr, warning=FALSE, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=7.5, fig.height=4.2, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.show = TRUE,
  hiline = TRUE,
  require("knitr")
)

```

```{r, echo=FALSE, warning=FALSE, include=FALSE}
library(survival)
library(tidyverse)
library(ggsurvfit)
library(lubridate)
library(ggfortify)

library(pscl)

library(broom)
library(broom.mixed)

library(readr)
```

```{r}
term <- 180
```

## Compare pre/post ibis for patients in first `r term` days in each period.

-   probability of inpatient admission

-   Mcnemar's test for paired proportions

-   \## stratify above by HCC score quantiles.

```{r}
event_file <- "../data/2025_02_18_utilization_events.csv"
report_file <- "../data/2025_02_18_utilization_report.csv"

source("read_clean.R")
```

```{r}
events_full %>% filter(org_name == "Unicare - Study") %>% summarize(mean(post_ibis_inpatient))

events_full %>% filter(org_name == "Unicare - Study") %>% 
   summarize(mean(event_start_date > ibis_coverage_end_date))

events_full %>% filter(org_name == "Unicare - Study") %>% 
  filter(event_start_date > ibis_coverage_end_date) %>% 
    summarize(mean(ibis_er))

```

```{r}
events_full %>% filter(org_name == "Unicare - Study") %>% 
    summarize(sum(pre_ibis_inpatient))
events_full %>% filter(org_name == "Unicare - Study") %>% 
    summarize(sum(ibis_inpatient))
```

```{r}
events_full %>% filter(org_name == "Unicare - Study") %>% 
    filter(event_start_date > ibis_coverage_start_date &
          event_start_date < ibis_coverage_end_date) %>% 
    summarize(sum(event_type == "inpatient"))
    
```

```{r}

```

```{r echo = FALSE}

ibis_inpatient_table <- events_full %>%  filter(org_name == "Unicare - Study")  %>% filter(ibis_inpatient == 1 | event_type == "no_utilization") %>% 
    group_by(pid, event_type)  %>%
    summarize(count = n(), .groups = "drop") %>% pivot_wider(names_from = event_type, values_from = count) %>%
    replace(is.na(.), 0) %>% rename(ibis_inpatient_count = inpatient)

```

```{r}

pre_ibis_inpatient_table <- events_full %>%  filter(org_name == "Unicare - Study")  %>% filter(pre_ibis_inpatient == 1 | event_type == "no_utilization") %>% 
    group_by(pid, event_type)  %>%
    summarize(count = n(), .groups = "drop") %>% pivot_wider(names_from = event_type, values_from = count) %>%
    replace(is.na(.), 0) %>% rename(pre_ibis_inpatient_count = inpatient)
```

```{r}
report_full %>% filter(org_name == "Unicare - Study") %>% 
    select(pid, total_ibis_inpatient) %>% head()

ibis_inpatient_table %>% filter(pid == 10473)
```

```{r}
ibis_inpatient_table %>% count(ibis_inpatient_count)
```

```{r}
inpatient_table <- pre_ibis_inpatient_table %>% full_join(ibis_inpatient_table, by = c("pid", "no_utilization")) %>% 
    mutate(ibis_inpatient_count = if_else(is.na(ibis_inpatient_count), 0, ibis_inpatient_count ))
```

```{r}

ibis_yes_no <- inpatient_table %>% 
    pivot_longer(c(3,4), names_to = "pre_post", values_to = "admit_count") %>% 
    mutate(admit = if_else(admit_count == 0, 0, 1))
```

```{r}
#ibis_inpatient_count 
```

```{r}
ibis_yes_no
```

```{r}
# pre- and post-IBIS inpatient counts
pre_ibis_counts <- report_full %>%
  filter(org_name == "Unicare - Study") %>%
  summarize(sum(pre_ibis_inpatient))

post_ibis_counts <- report_full %>%
  filter(org_name == "Unicare - Study") %>%
  summarize(sum(post_ibis_inpatient))

```

## Wilcox - does it assume normality

```{r}
# look at admissions using total_ibis_inpatient
admit_table <- report_full %>%
  filter(org_name == "Unicare - Study") %>%
  select(total_pre_ibis_inpatient, total_ibis_inpatient) 

admit_table
table(admit_table$total_pre_ibis_inpatient, admit_table$total_ibis_inpatient)
```

# check this versus the difference

# next take out patients with zero days on ibis

# look at the counts per number of days coverage pre/post

```{r}
#wilcox is paired 

#Wilcoxon test
wilcox_test_result <- wilcox.test(admit_table$total_pre_ibis_inpatient, 
                                  admit_table$total_ibis_inpatient, 
                                  data = report_full, 
                                  paired = TRUE)

print(wilcox_test_result)
```

```{r}
# filter coverage days >0
admit_table_filtered <- report_full %>%
  filter(org_name == "Unicare - Study") %>%
  filter(covered_days_pre_ibis > 0) %>%
  filter(covered_days_on_ibis > 0) %>%
select(total_pre_ibis_inpatient, total_ibis_inpatient) 

# Wilcoxon test: Pre vs Post IBIS days
wilcoxon_result <- wilcox.test(
  admit_table_filtered$total_pre_ibis_inpatient,
  admit_table_filtered$total_ibis_inpatient,
  paired = TRUE
)

print(admit_table_filtered)
print(wilcoxon_result)
```

# Next thing -- filtering out patients that had a certain number of days in each period, filtering only the events that happened

inpatient +/- term ibis_cov_start + 180 or - 180

```{r}
# count only events
#pre_term <- events_full %>% 
#  filter(ibis_coverage_start_date - 180 days # so 180 days before start date)
         
#prost_term <- events_full %>% 
#  filter(ibis_coverage_start_date + 180 days # so 180 days after start date)


term <- 180

ibis_term <- report_full %>%
    filter(org_name == "Unicare - Study") %>%
    filter(ibis_coverage_end_date - ibis_coverage_start_date > term) %>%
    filter(ibis_coverage_start_date - coverage_start_date > term) %>%
    mutate(days = ibis_coverage_end_date - ibis_coverage_start_date)

events_during_ibis_term <- ibis_term %>%
  group_by(pid, days) %>%
  summarize(inpatient_count = sum(event_type == "inpatient"))

print(events_during_ibis_term)

```

```{r}
ibis_term <- report_full %>%
    filter(org_name == "Unicare - Study") %>%
    filter(ibis_coverage_end_date - ibis_coverage_start_date > term) %>%
    filter(ibis_coverage_start_date - coverage_start_date > term) %>%
    mutate(days = ibis_coverage_end_date - ibis_coverage_start_date)

events_during_ibis_term <- ibis_term %>%
  group_by(pid, days) %>%
  summarize(
    pre_ibis_inpatient_count = sum(event_type == "inpatient" & coverage_start_date < ibis_coverage_start_date),
    post_ibis_inpatient_count = sum(event_type == "inpatient" & coverage_end_date > ibis_coverage_end_date)
  )

print(events_during_ibis_term)



```

```{r}
pre_ibis <- report_full %>%
    filter(org_name == "Unicare - Study") %>%
    filter(ibis_coverage_start_date - coverage_start_date > term) %>%
    filter(coverage_start_date >= (ibis_coverage_start_date - 180)) %>%
    group_by(pid) %>%
    summarize(pre_ibis_inpatient_count = sum(event_type == "inpatient"))
pre_ibis
```

```{r}
events_during_ibis_term <- ibis_term %>%
  filter(days >= 180) %>%  # Ensures at least 180 days of coverage
  group_by(pid, days) %>%
  summarize(
    pre_ibis_inpatient_count = sum(event_type == "inpatient" & coverage_start_date < ibis_coverage_start_date),
    post_ibis_inpatient_count = sum(event_type == "inpatient" & coverage_end_date > ibis_coverage_end_date)
  )

print(events_during_ibis_term)

```
